---
title: "HW2 Q2"
author: "Danai Avgerinou, Holly Capell, Shannon McNish, Taylor Pellerin, Kaya Tollas"
date: "4/22/2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(igraph)
library(sand)
library(tidyverse)
```


```{r}
el <- read_csv("~/CompStats/HW/hw_2/weighted_edges.csv")
el = as.matrix(el)
marvel <- graph.data.frame(el[,1:2], directed = F)
E(marvel)$weight <- as.numeric(el[,3])
```

# Avenger analysis

Below we investigate the universe of the Avengers as compared to Captain America's neighborhood. This will give us insight into the relative importance of the Avengers 

```{r}
marvel_nodes <-  V(marvel)
captain_america <- which(marvel_nodes$name == 'CAPTAIN AMERICA')
thor <- which(marvel_nodes$name == 'THOR/DR. DONALD BLAK')
iron_man <- which(marvel_nodes$name == 'IRON MAN/TONY STARK')
black_panther <- which(marvel_nodes$name == 'BLACK PANTHER/T\'CHAL')
captain_marvel <- which(marvel_nodes$name == 'CAPTAIN MARVEL II/MO')
ghost_rider <- which(marvel_nodes$name == 'GHOST RIDER II/JOHNN')
she_hulk <- which(marvel_nodes$name == 'SHE-HULK/JENNIFER WA')
hulk <- which(marvel_nodes$name == 'HULK/DR. ROBERT BRUC')
black_widow <- which(marvel_nodes$name == 'BLACK WIDOW/NATASHA')
dr_strange <- which(marvel_nodes$name == 'DR. STRANGE/STEPHEN')
ant_man <- which(marvel_nodes$name == 'ANT-MAN/DR. HENRY J.')
hawk <- which(marvel_nodes$name == 'HAWK')

avenger_ids <- c(captain_america, thor, iron_man, black_panther, black_widow, captain_marvel, ghost_rider, she_hulk, hulk, dr_strange, hawk, ant_man)
avengers <-  induced.subgraph(marvel, avenger_ids)

plot(avengers, layout = layout.kamada.kawai, vertex.size = degree(avengers), edge.width = log( E(avengers)$weight))
```
Now we investigate betweenness centrality of the Avengers. Suprisingly, if we remove Captain America, all the Avengers are still connected. The most between-hero in the (selected) Avengers is actually Ghost Rider!

```{r}
betweenness(avengers) %>% sort(decreasing=T)
```
Now we look at eigenvector centrality, which shows that within the Avengers, Iron Man is the next-most important in terms of his important friends.

```{r}
eigen_centrality(avengers)$vector %>% sort(decreasing=T)
```
Now we will visualize Captain America's ego network. This is a subgraph of popular heroes connected to Captain America who are also highly-connected in the network. The red node is Captain America and the yellow ones are Avengers. All other nodes are heroes from different leagues. 

```{r }
ego.captain <- induced.subgraph(marvel, igraph::neighborhood(marvel, 1, 'CAPTAIN AMERICA')[[1]])
vertices.captain <-  V(ego.captain)
friends <- degree(ego.captain) %>% sort(decreasing=T)
irrelevant <- degree(ego.captain) < friends[20] #only keep his most connected friends
ego.cap <- delete_vertices(ego.captain, vertices.captain[irrelevant])
vertices.captain <-  V(ego.cap)

captain_america <- which(vertices.captain$name == 'CAPTAIN AMERICA')
thor <- which(vertices.captain$name == 'THOR/DR. DONALD BLAK')
iron_man <- which(vertices.captain$name == 'IRON MAN/TONY STARK')
black_widow <- which(vertices.captain$name == 'BLACK WIDOW/NATASHA')
black_panther <- which(vertices.captain$name == 'BLACK PANTHER/T\'CHAL')
captain_marvel <- which(vertices.captain$name == 'CAPTAIN MARVEL II/MO')
ghost_rider <- which(vertices.captain$name == 'GHOST RIDER II/JOHNN')
she_hulk <- which(vertices.captain$name == 'SHE-HULK/JENNIFER WA')
hulk <- which(vertices.captain$name == 'HULK/DR. ROBERT BRUC')
black_widow <- which(vertices.captain$name == 'BLACK WIDOW/NATASHA')
dr_strange <- which(vertices.captain$name == 'DR. STRANGE/STEPHEN')
ant_man <- which(vertices.captain$name == 'ANT-MAN/DR. HENRY J.')
hawk <- which(vertices.captain$name == 'HAWK')

avenger_ids <- c(captain_america, thor, iron_man, black_panther, hulk, black_widow, captain_marvel, ghost_rider, she_hulk, dr_strange, ant_man, hawk)

hero_type <- rep(1,vcount(ego.cap))
hero_type[avenger_ids] <- 2
hero_type[captain_america] <- 3

# specify a vector of 5 colors for other, Avengers, and Captain America
colorvec <- c('black','yellow', 'red')
# assign colors to nodes
V(ego.cap)$color <- colorvec[hero_type]
plot(ego.cap, layout = layout.kamada.kawai, vertex.label = NA)
```

Now we check closeness centrality for Captain America's co-hero network. 

```{r}
closeness(ego.cap, mode = c("out", "in", "all", "total"),
  weights = NULL, normalized = FALSE) %>% sort(decreasing=T)
```
It appears everyone is approximately equidistant on this network.

Now we look at eigenvector centrality. It is interesting to note that after Captain America, the most central nodes are all part of the Fantastic Four. A few X-Men show up here too, which confirms that those leagues are the most popular and influential in the MCU. 

```{r}
eigenvector.centrality <- eigen_centrality(ego.cap, directed = FALSE)$vector
eigenvector.centrality %>% sort(decreasing=T)
```


